Anomaly detection with kernel density estimation on manifolds

### 2022-05-05 paper/paper.Rmd
True volume density function to be used, also try small N, optimize h with AMSE

Let $S^2$ be the two-dimensional sphere of radius 1 and $p \in S^2$. Let
$v$ and $w$ be vectors such that $\|v\| = \|w\| = 1$ and $\{v, w, p\}$
is the orthonormal basis of $\mathbb{R}^3$. Consider the exponential
chart $(U, \Psi)$ induced by the parametrisation
$\Psi^{-1}: B_\pi(0) \rightarrow S^2 \backslash \{-p\}$ given by $$
\Psi^{-1}(s,t) = \cos(r)p + \sin(r)(sv+tw)r^{-1},
$$ if $(s,t) \neq (0,0)$ and $\Psi^{-1}(0,0)=p$ where
$r=\sqrt{s^2 + t^2}$. Note that $r=d_g(p,q)$ if $q = \exp_p(sv+tw)$.
Then we have $$
A(s,t) = \frac{\partial}{\partial \Psi_1} \bigg|_{\Psi^{-1}(s,t)}, \\
B(s,t) = \frac{\partial}{\partial \Psi_2} \bigg|_{\Psi^{-1}(s,t)}.
$$ Then each component of the riemannian metric could be calculated as
$g(A(s,t), A(s,t)), g(A(s,t), B(s,t)), g(B(s,t), B(s,t))$. We obtain
that the volume density on the sphere is $$
\theta_p(q) = \bigg| \det g_q\bigg( \frac{\partial}{\partial \Psi_i}\bigg|_q, \frac{\partial}{\partial \Psi_j}\bigg|_q \bigg) \bigg|^{1/2} \\
= \frac{|\sin(d_g(p,q))|}{d_g(p,g)},
$$ for $q \neq p, -p$, and $\theta_p(p) = 1$.

In $S^1$, polar coordinates: $\phi(\exp_p{vt})^{\prime}(0) = v$

cartesian coordinates:
$\psi(\exp_p{vt})^{\prime}(0) = x_p \cos{v} - \sqrt{1-x_p^2} \sin{v}$
\bigg\_t=0 = -\sqrt{1-x_p^2}v\$

>>>>

Let $S^2$ be the two-dimensional sphere of radius 1 and $p \in S^2$. Let
$v$ and $w$ be vectors such that $\|v\| = \|w\| = 1$ and $\{v, w, p\}$
is the orthonormal basis of $\mathbb{R}^3$. Consider the exponential
chart $(U, \Psi)$ induced by the parametrisation
$\Psi^{-1}: B_\pi(0) \rightarrow S^2 \backslash \{-p\}$ given by $$
\Psi^{-1}(s,t) = \cos(r)p + \sin(r)(sv+tw)r^{-1},
$$ if $(s,t) \neq (0,0)$ and $\Psi^{-1}(0,0)=p$ where
$r=\sqrt{s^2 + t^2}$. Note that $r=d_g(p,q)$ if $q = \exp_p(sv+tw)$.
Then we take the partial derivatives of \$\\Psi\^{-1}(s,t)\$ with
respect to \$s\$ and \$t\$ and have that$$
A(s,t) = \frac{\partial}{\partial \Psi_1} \bigg|_{\Psi^{-1}(s,t)} = \frac{-\sin{(r)}s}{r}p + \frac{\cos(r)rs^2+t^2\sin(r)}{r^3}v + \frac{(\cos(r)r-\sin(r))st}{r^3}w, \\
B(s,t) = \frac{\partial}{\partial \Psi_2} \bigg|_{\Psi^{-1}(s,t)} = \frac{-\sin{(r)}t}{r}p + \frac{(\cos(r)r-\sin(r))st}{r^3}v + \frac{\cos(r)rt^2+\sin(r)s^2}{r^3}w.
$$ Then each component of the riemannian metric could be calculated as
$g(A(s,t), A(s,t)), g(A(s,t), B(s,t)), g(B(s,t), B(s,t))$. We obtain
that the volume density on the sphere is $$
\theta_p(q) = \bigg| \det g_q\bigg( \frac{\partial}{\partial \Psi_i}\bigg|_q, \frac{\partial}{\partial \Psi_j}\bigg|_q \bigg) \bigg|^{1/2} \\
= \frac{|\sin(d_g(p,q))|}{d_g(p,g)},
$$ for $q \neq p, -p$, and $\theta_p(p) = 1$.

In $S^1$, we could let \$\\{v, p\\}\$ be the basis of
\$\\mathbb{R}\^2\$. Then the exponential chart induced

Polar coordinates: $\phi(\exp_p{vt})^{\prime}(0) = v$

Cartesian coordinates:
$\psi(\exp_p{vt})^{\prime}(0) = (x_p \cos{v} - \sqrt{1-x_p^2} \sin{v})^\prime  \bigg|_{t=0} = - \sqrt{1 - x_p^2}v$
